---
title: "Midterm Project"
author: "Yihong Hu, Sabir Nazarov, Will Friedrichs"
date: "10/20/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    runtime: shiny
# >>>>>>> 05ab802ca33d750c6faefee82b91766e0f572e3e
---

# Introduction

Zillow's housing market predictions are an integral part of its business model, helping the company achieve a greater understanding of how the market will value properties. Our team is confident that through our geospatial machine learning-based model that considers not only attributes of homes, but local factors as well, we can improve Zillow's house price predictions for the Boulder County study area and provide a template which can be adapted to other localities.

This project outlines and analyzes the process by which we created our model in four broad stages: data gathering, feature creation, multivariate regression modeling, and prediction generation. We present these stages sequentially, though the ordering does not necessarily reflect the order in which parts of the model were constructed.

Along with written analysis, plots, and tables, the R code that produces the model is presented in the document in chunks. The chunk below, for example, loads 17 libraries for use later, helps set up the generation of the r markdown file, gives access to online resources helpful to our analysis, and sets up a color scheme to be used for data visualization.

```{r The Setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(dplyr)
library(ggcorrplot)
library(caret)
library(spdep)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(jtools)     
library(ggstance)
library(rpart)
library(ggplot2)
library(stargazer)

# set up knitr
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	cache = TRUE,
	echo=TRUE
)

# Set root directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

# Locate the source of functions from raw.githubusercontent.com
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

# Save q5 and qbr functions for later use
q5 <- function(variable) {as.factor(ntile(variable, 5))}

qbr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

# Set a color palette for later use
palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```

The data gathering process begins by accessing tract-level geography, population, and demographic data from the US Government Census API. The data uses American Community Survey results for 2019. Variables of interest are isolated and wrangled to create features required for the analysis.


```{r ACS Data, message=FALSE, warning=FALSE}
varlist_2019 <- load_variables(2019, "acs5", cache = TRUE)

census_api_key("94efffd19b56ad527e379faea1653ee74dc3de4a",overwrite = TRUE)

tracts19 <- get_acs(geography = "tract",                         
                    variables = c("B01001_001","B23025_004","B19001_001",
                                  "B06012_002","B02001_002","B25002_003",
                                  "B25013_006","B08013_001","B15012_009"), 
                    year=2019, 
                    state=08, 
                    county=013,
                    output = "wide",
                    geometry=TRUE) %>% 
            st_transform('ESRI:102254') %>%
            select( c("GEOID","B01001_001E","B23025_004E","B06011_001E",
                      "B06012_002E","B02001_002E","B25002_003E",
                      "B25013_006E","B08013_001E","B15012_009E","geometry") ) %>%
            rename(tot_pop = "B01001_001E",
                   empl_pop = "B23025_004E",
                   med_inc = "B19001_001",
                   pvty_pop = "B06012_002E",
                   white_pop = "B02001_002E",
                   vac_occ = "B25002_003E",
                   own_occ_bach = "B25013_006E",
                   tt_work = "B08013_001E",
                   sci_bach = "B15012_009E") %>%
            mutate(area = as.numeric(st_area(geometry)/1000000))%>%
            mutate(pop_den = tot_pop/area)
```


## Project Data

The primary training dataset provides information on a variety of features associated with houses, along with house prices. Some of the houses have prices listed, while others do not. After other variables are added to the dataset for each house, OLS regression uses houses with price information to create our model, which is a formula that can produce price from those oteher variables. 

Given the original primary dataset, we wrangled and joined ACS tract-based demographic data that we theorized could predict housing prices. Each house is ascribed the demographic variables associated with its tract.

Boulder County and the boundaries of the municipalities within are established from open-source geometry files. Table 1, below, summarizes key indicators by municipality. Each variable represents the mean values ascribed to homes in the municipality. For example, total population represents not the population of the municipality, but the mean tract population ascribed to each house. The table reveals that the price is strongly related to municipality. 


```{r Project Data, message=FALSE, warning=FALSE, echo=FALSE}
# This loads all the data into "studentData".
studentData <- st_read("studentData.geojson", crs = 'ESRI:102254')%>%
  mutate(Age = 2021 - builtYear) %>%
  mutate(TotalBath = nbrThreeQtrBaths + nbrFullBaths + nbrHalfBaths)

studentData %>%
st_make_valid(geometry)

# Attach ACS data
studentData <- st_join(studentData, tracts19, join = st_within)

#Boulder County Boundary
BoulderCounty_Bundary <-
st_read("https://opendata.arcgis.com/datasets/964b8f3b3dbe401bb28d49ac93d29dc4_0.geojson")%>%
  select(geometry) %>%
  st_transform('ESRI:102254')

#Boulder Municipal Boundary
BoulderMuni_Boundary <-
  st_read("https://opendata.arcgis.com/datasets/9597d3916aba47e887ca563d5ac15938_0.geojson")%>%
  st_transform('ESRI:102254')%>%
  rename(Municipality = ZONEDESC)
  

#Map of Boulder County with Housing Sales Price and municipal boundary
ggplot()+
  geom_sf(data = BoulderCounty_Bundary, fill = "grey70") +
  geom_sf(data = BoulderMuni_Boundary, aes(fill = Municipality, alpha=0.5),colour = "white") +
  geom_sf(data = studentData, aes(colour = q5(price)),size=.85)+
  scale_colour_manual(values = palette5,
                      labels=qbr(studentData,"price"),
                      name = "Sale Price") 

# Median Housing Price in Each Municipality
House_in_muni_boundary <- st_intersection(studentData, BoulderMuni_Boundary) %>%
  mutate(pct_pvty = pvty_pop/tot_pop * 100,
         pct_white = white_pop/tot_pop * 100,
         )

Municipality.Summary <-
  st_drop_geometry(House_in_muni_boundary) %>%
  group_by(Municipality) %>%
  summarize(Medium.Price = median(price, na.rm = T),
            Mean.Price = mean(price, na.rm = T),
            Population = mean(tot_pop, na.rm = T),
            "Poplation Density per km^2"= mean(pop_den,na.rm=T),
            "Median Income" = mean(med_inc, na.rm = T),
            "Poverty Rate" = mean(pct_pvty, na.rm = T),
             "White Poplation in %" = mean(pct_white, na.rm = T),
            "Building Age" = mean(Age,na.rm = T)) %>%
  arrange(desc(Medium.Price))

kable(Municipality.Summary, digits = 2) %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 1")


studentData <- st_join(studentData, BoulderMuni_Boundary, join = st_within)

studentData <- 
studentData %>%
  select(-c(OBJECTID, ZONECLASS, LASTUPDATE,LASTEDITOR, REG_PDF_URL, Shape_STArea__, 
            Shape_STLength__, ShapeSTArea,ShapeSTLength
            ))

# Creating dummy variables for municipalities
studentData <- mutate(studentData, Loui_dummy = case_when(Municipality=="Louisville"~ 1, Municipality!="Louisville"~ 0))
studentData <- mutate(studentData, Ward_dummy = case_when(Municipality=="Ward"~ 1, Municipality!="Ward"~ 0))
studentData <- mutate(studentData, Jame_dummy = case_when(Municipality=="Jamestown"~ 1, Municipality!="Jamestown"~ 0))
studentData <- mutate(studentData, Nede_dummy = case_when(Municipality=="Nederland"~ 1, Municipality!="Nederland"~ 0))
studentData <- mutate(studentData, Boul_dummy = case_when(Municipality=="Boulder"~ 1, Municipality!="Boulder"~ 0))
studentData <- mutate(studentData, Erie_dummy = case_when(Municipality=="Erie"~ 1, Municipality!="Erie"~ 0))
studentData <- mutate(studentData, Lafa_dummy = case_when(Municipality=="Lafayette"~ 1, Municipality!="Lafayette"~ 0))
studentData <- mutate(studentData, Long_dummy = case_when(Municipality=="Longmont"~ 1, Municipality!="Longmont"~ 0))
studentData <- mutate(studentData, Lyon_dummy = case_when(Municipality=="Lyons"~ 1, Municipality!="Lyons"~ 0))
studentData <- mutate(studentData, Supe_dummy = case_when(Municipality=="Superior"~ 1, Municipality!="Superior"~ 0))

studentData$Loui_dummy[is.na(studentData$Loui_dummy)] <- 0
studentData$Ward_dummy[is.na(studentData$Ward_dummy)] <- 0
studentData$Jame_dummy[is.na(studentData$Jame_dummy)] <- 0
studentData$Nede_dummy[is.na(studentData$Nede_dummy)] <- 0
studentData$Boul_dummy[is.na(studentData$Boul_dummy)] <- 0
studentData$Erie_dummy[is.na(studentData$Erie_dummy)] <- 0
studentData$Lafa_dummy[is.na(studentData$Lafa_dummy)] <- 0
studentData$Long_dummy[is.na(studentData$Long_dummy)] <- 0
studentData$Lyon_dummy[is.na(studentData$Lyon_dummy)] <- 0
studentData$Supe_dummy[is.na(studentData$Supe_dummy)] <- 0
```

### Parks, Landmarks, and Trailheads

Considering that Boulder region is rich in natural beauty and outdoor activities, we use the buyers logic that places a premium on proximity to these locations.  We incorporate the following data into the analysis: parks, natural landmarks, and entrance to trail heads.  For example, we create a variable that captures the number of parks within a 500 meter radius, trail heads within 1000 meter radius.


```{r Parks & Landmarks}
# Load Park data
GreenSpacePolygon <- st_read("County_Open_Space.geojson") %>%
                     st_transform('ESRI:102254')

Park <- GreenSpacePolygon[!is.na(GreenSpacePolygon$PARK_GROUP),] %>%
        st_centroid()
  
st_c <- st_coordinates


studentData <-
  studentData %>% 
  mutate(park_nn1 = nn_function(st_c(studentData), st_c(Park), 1),
         park_nn2 = nn_function(st_c(studentData), st_c(Park), 2),
         park_nn3 = nn_function(st_c(studentData), st_c(Park), 3),
         park_dist = st_distance(studentData,Park))

Park_in_buffer <-
  st_join(Park,st_buffer(studentData,500),join = st_within)

PKCount <-
  Park_in_buffer %>%
  count(MUSA_ID)%>%
  st_drop_geometry()

studentData <-
  left_join(studentData, PKCount, by = "MUSA_ID" )

studentData <-
  studentData %>%
  rename(PKCount500m = n)

studentData$PKCount500m[is.na(studentData$PKCount500m)] <- 0

# attach distance to green space data
# studentData %>% mutate(green_dis = st_distance(studentData, GreenSpacePolygon))

# Load Landmarks data
landmarksPolygon <- st_union(st_read("Natural_Landmarks.geojson")) %>%
  st_transform('ESRI:102254')

# attach distance to land marks data
studentData <- mutate(studentData, landmark_dist = st_distance(studentData, landmarksPolygon))

# Loading Trail Heads Locations
TrailHead <- 
  st_read("https://opendata.arcgis.com/datasets/5ade4ef915c54430a32026bcb03fe1d7_0.geojson") %>%
  st_transform('ESRI:102254')%>%
  select(geometry)

#Apply buffer counts and nearest neighbor function on trail heads
Trailhead_in_buffer <-
  st_join(TrailHead,st_buffer(studentData,1000),join = st_within)

TrailHeadCount <-
  Trailhead_in_buffer %>%
  count(MUSA_ID)%>%
  st_drop_geometry()

studentData <-
  left_join(studentData, TrailHeadCount, by = "MUSA_ID" )

studentData <-
  studentData %>%
  rename(TrailHead1000m = n)

studentData$TrailHead1000m[is.na(studentData$TrailHead1000m)] <- 0

studentData <-
  studentData %>% 
  mutate(
    head_nn1 = nn_function(st_c(studentData), st_c(TrailHead), 1),
    head_nn2 = nn_function(st_c(studentData), st_c(TrailHead), 2), 
    head_nn3 = nn_function(st_c(studentData), st_c(TrailHead), 3),
    head_nn4 = nn_function(st_c(studentData), st_c(TrailHead), 4), 
    head_nn5 = nn_function(st_c(studentData), st_c(TrailHead), 5))

studentData <-
  studentData %>%
  mutate(trail_dist = st_distance(studentData, TrailHead))
```


### Playgrounds, Schools, Flood Plains

Amenities and public services are important aspects to choose one's location for his/her house. Considering most of the house buyers are families with children, playgrounds and schools are chosen as indicators: we create variables that captures the number of playgrounds and schools within a 500 meter radius. Housing distance to flood plains are considered due to concern of safety. Boulder County has encountered high rate of flooding for the past decade.

```{r PLaygrounds & Schools}
#Loading Playground Locations
Playground <- 
  st_read("Playground_Sites_Points.GEOJSON") %>%
  st_transform('ESRI:102254')%>%
  drop_na(PROPID)%>%
  select(geometry)

Playground.sf <-
  st_join(Playground,st_buffer(studentData,500),join = st_within)

PGCount <-
  Playground.sf %>%
  count(MUSA_ID)%>%
  st_drop_geometry()

studentData <-
  left_join(studentData, PGCount, by = "MUSA_ID" )
  
studentData <-
  studentData %>%
  rename(pgcount500m = n)

studentData$pgcount500m[is.na(studentData$pgcount500m)] <- 0
  

#Loading School Locations
Schools <- 
  st_read("CDPHE_CDOE_School_Locations_and_District_Office_Locations.GEOJSON")%>%
  st_transform('ESRI:102254')%>%
  filter(COUNTY == "BOULDER")
  
Private_School <-
  filter(Schools, startsWith(Type_, "Non-"))

studentData <-
  studentData %>% 
  mutate(
  school_nn1 = nn_function(st_c(studentData), st_c(Schools), 1))

studentData <-
  studentData %>%
  mutate(
    privatesch_nn1 = nn_function(st_c(studentData),st_c(Private_School),1),
    privatesch_nn2 = nn_function(st_c(studentData),st_c(Private_School),2),
    privatesch_nn3 = nn_function(st_c(studentData),st_c(Private_School),3),
    privat_dist = st_distance(studentData,Private_School, by_element = TRUE))

#Loading Flood Plain
Floodplain <-
 st_read("https://opendata.arcgis.com/datasets/30674682e55e4e1b8b0e407b0fe23b9a_0.geojson") %>%
 st_transform('ESRI:102254')%>%
  st_union()

studentData <-
  studentData %>%
  mutate(flood_dist = st_distance(studentData, Floodplain))


# This selects all numeric variables as preparation for the
# correlation analysis that follows.
#(also delete the outliar)

cleanData <- 
  select_if(st_drop_geometry(studentData), is.numeric) %>%
  select(!c(year, ExtWallSec, IntWall, Roof_Cover, Stories, UnitCount, MUSA_ID))%>%
  slice(-2638)

cleanData$Ac[(cleanData$Ac)== 0] <- NA
cleanData$Heating[(cleanData$Heating) == 0] <- NA
```

# The Model

## Preparation for machine learning and regression model

Regression model is a good predicting method that considers all the interested variables. In order to prepare for the regression, we need to train our data by machine learning and cross-validate our to see if the model is appropriate to generalize on other data sets. 

First, We do a 75/25 split of the data with price information with selected variables. 75% of the data is to be trained on. 25% is to be tested on for validation. 

```{r Train & Test}
# The test data only includes rows comprising the test set.
# Remove Outliars

test.Data <- filter(cleanData, toPredict == 1)

#Split the ‘toPredict’ == 0 into a separate training and test set 
#using a 75/25 split

train.Data <- filter(cleanData, toPredict == 0) %>%
  na.omit()

inTrain <- createDataPartition(
  y = paste(train.Data$Heating, train.Data$Ac), 
  p = .75, list = FALSE)
train.Data.1 <- train.Data[inTrain,] 
train.Data.2 <- train.Data[-inTrain,]

# This function here attempts to fit a linear model to
# the relationship between "testData" variables.
ggplot(data = train.Data, aes(Heating, price)) +
       geom_point(size = .5) + 
       geom_smooth(method = "lm")

#Dataframe with variables of interest
train.Data.clean <-
  train.Data %>%
  select(price, 
         section_num,
         qualityCode,
         TotalFinishedSF,
         Age,
         Ac,
         Heating,
         med_inc,
         nbrRoomsNobath,
         vac_occ,
         tot_pop,
         pop_den,
         pvty_pop,
         head_nn5,
         park_nn1,
         Boul_dummy,
         Loui_dummy,
         pgcount500m,
         TotalBath,
         flood_dist,
         landmark_dist,
         privat_dist)
```


## Correllation & Test Signifcance

Prior to the the regression set-up, the relationship across all the variables are evaluated to see if they are significant indicator for prediction. In Fig.1 Pearson Correlation is imputed: a strong orange shows strong positive correlation and a strong green shows strong negative correlate between selected variables. 

Note that price is strongly correlated with section number, quality code, total finished square footage, number of bathrooms, if it's the city of Boulder, and count of playground within 500 meter radius. 

Fig 2 is a in-depth look of Fig 1 where we evaluate correlation between price and selected variable. Strong positive correlation is seen between price and total square footage. Negative correlation is seen between price and distance away from landmark and private schools, indicating the further away from landmark and private school, the less the price.

```{r Correllation & Test Signicance, class.source = 'fold-show', echo=FALSE}

# Correlation analysis: pearson for each relationship
ggcorrplot(
  round(cor(train.Data.clean), 1), 
  p.mat = cor_pmat(train.Data.clean),show.diag = TRUE,
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
  labs(title = "Correlation across numeric variables",caption = "Fig 1") 

#Scatter plots between price and variables
cleanData %>%
  select(price,"Playground Count in 500m" = pgcount500m, 
         "Total Squrefootage" = TotalFinishedSF,"Distance to Landmark (m)" = landmark_dist,
         "Distance to the nearest private school" = privat_dist)%>%
gather(Variable, Value, -price) %>% 
  ggplot(aes(Value, price)) +
  geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
  facet_wrap(~Variable, ncol = 3, scales = "free") +
  labs(title = "Price as a function of continuous variables", caption = "Fig 2") +
  plotTheme()

# This function allows for the plug-in of variables from "studentData".
testSignificance <- lm(price ~ ., data = train.Data %>% 
                    dplyr::select(price, 
                                  qualityCode,
                                  TotalFinishedSF,
                                  mainfloorSF,
                                  Age,
                                  pop_den
                                  ))

# This gives us our r-squared value, which measures fit to the training data.
summary(testSignificance)

```


## K-folds & Multivariate

K-folds validates the regression model. We divide the data into 100 folds, and run regression over each fold to compare the errors between observed price and the predicted price. 

Table 1. returns the result that tells us if how well each variable performs in the prediction: P-value < 0.05 indicates the variable is siginificant. R-square at the end gives a percentage of how well our model can control the error. 


```{r}
# This sets up k-fold cross-validation.
k = 100
fitControl <- trainControl(method = "cv", number = k)
set.seed(825)

# Multivariate regression 
reg1 <- lm(price ~ ., data = train.Data.1 %>% 
             dplyr::select(price, 
                           section_num,
                           qualityCode,
                           TotalFinishedSF,
                           Age,
                           Ac,
                           Heating,
                           med_inc,
                           nbrRoomsNobath,
                           landmark_dist,
                           white_pop,
                           pop_den,
                           pvty_pop,
                           privat_dist,
                           head_nn5,
                           park_nn1,
                           Loui_dummy,
                           Ward_dummy,
                           Jame_dummy,
                           Nede_dummy,
                           Boul_dummy,
                           Erie_dummy,
                           Lafa_dummy,
                           Long_dummy,
                           TotalBath,
                           pgcount500m,
                           flood_dist))

stargazer(reg1, type = "html", title = "Table 1. Variation of selected variable and Error Control of the regression model",out="table1.txt",dep.var.labels = "Housing Price")

reg1.tidy <-
  tidy(reg1,
       "Section Number" = section_num)

kable(reg1.tidy, digits = 2, caption = "Table 2. Regression Model Predicting Variation on Housing Sales Price", 
      col.names = c("Predictor","Estimate","Standard Error", "T-Value","P-Value")
      ) %>%
  kable_styling()

# variables in the "select(...)" function are considered in the analysis here.
regression.100foldcv <- 
  train(price ~ ., data = train.Data.1%>% 
          select(price, 
                 section_num,
                 qualityCode,
                 TotalFinishedSF,
                 Ac,
                 Heating,
                 Age,
                 med_inc,
                 landmark_dist,
                 nbrRoomsNobath,
                 pvty_pop,
                 privat_dist,
                 head_nn5,
                 park_nn1,
                 Loui_dummy,
                 Nede_dummy,
                 Boul_dummy,
                 Erie_dummy,
                 Long_dummy,
                 Lyon_dummy,
                 pgcount500m,
                 TotalBath,
                 flood_dist,
                 privat_dist
          ),
  method = "lm", trControl = fitControl, na.action = na.pass)


# The resulting Mean Absolute Error (MAE) of running this line tells us how
# successful our model is at predicting unknown data. 
regression.100foldcv

```


## MAE Histogram

Describe what we did here.

```{r MAE Histogram}
#################
#MAE Histogram
#This allows you to see the distribution of MAE of the folds

ggplot(regression.100foldcv$resample, aes(x=MAE))+
  geom_histogram(colour = "white",fill="orange")+
  labs(title="Distribution of MAE",x="Mean Absolute Error", y = "Count",
       subtitle = "k-fold cross-validation; k = 100",
       caption = "Figure 2")+
  theme_apa()
```


## Predictions

Describe what we did here.

```{r}
#################
#Run model on unpredicted dataset

train.Data.2 <-
  train.Data.2 %>%
  mutate(SalePrice.Predict = predict(reg1, train.Data.2))
         
train.Data.2 <-
  train.Data.2 %>%
        mutate(SalePrice.Error = SalePrice.Predict - price,
         SalePrice.AbsError = abs(SalePrice.Predict - price),
        SalePrice.APE = (abs(SalePrice.Predict - price)) / SalePrice.Predict)

#Table of mean absolute error and MAPE
Prediction.Variation <- "Trial 1" 
MeanAb <- mean(train.Data.2$SalePrice.AbsError,na.rm = T)
MeanAPE <- mean(train.Data.2$SalePrice.APE, na.rm = T)

MeanAb.table <- data.frame("Model"= Prediction.Variation,"Mean Absolute Error ($)"= MeanAb,
                           "Mean Absolute Percentage Error (%)" = MeanAPE)

kable(MeanAb.table, digit = 2,
      caption = "Table 3. Mean Absolute Error and Absolute
      Percentage Error on the Prediction Trial",
      col.names = c("Regression",
                    "Mean Absolute Error ($)",
                    "Mean Absolute Percentage Error (%)"))%>%
  kable_styling(latex_options = "HOLD_position")
```


## Plot Predictions

Describe what we did here.

```{r Plot Predictions}
#Plot of predicted prices as a function of obersved price
ggplot(data = train.Data.2, aes(price,SalePrice.Predict)) +
  geom_point(size = .85,colour = "orange") + 
  geom_smooth(method = "lm",colour = "red",size = 1.2)+
  labs(x = "Obeserved Prices",y = "Predicted Prices", title = "Predicted Prices as a Function of Observed Prices",
       subtitle = "Boulder County, CO (all in $)")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.line = element_line(colour = "grey50", size = 1),
        panel.grid.major = element_line(linetype = "dotted",size = 1))

```

## Lag Price Errors

Describe what we did here.

```{r Lag Price Errors}
##################
#Lag Price Errors

train.Data.2 <-
  left_join(train.Data.2,studentData)%>%
  st_as_sf()

coords.test <-  st_coordinates(train.Data.2) 
neighborList.test <- knn2nb(knearneigh(coords.test, 5))
spatialWeights.test <- nb2listw(neighborList.test, style="W")

train.Data.2.lag <-
train.Data.2 %>% 
  mutate(lagPriceError = lag.listw(spatialWeights.test, SalePrice.Error))

ggplot(data = train.Data.2.lag, aes(lagPriceError, SalePrice.Error)) +
  geom_point(size = .85,colour = "orange") + 
  geom_smooth(method = "lm",colour = "red",size = 1.2)+
  labs(x = "Spatial Lag of Errors (Mean error of five nearest neighbors)",y = "Sale Price Error", title = "Lag Sale Price Errors against Predicted Sale Price Error",
       subtitle = "Boulder County, CO (all in $)")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.line = element_line(colour = "grey50", size = 1),
        panel.grid.major = element_line(linetype = "dotted",size = 1))
```



## Moran's I

Describe what we did here.
```{r Morans I}
#Moran's I

moranTest <- moran.mc(train.Data.2$SalePrice.Error, 
                      spatialWeights.test, nsim = 999)

ggplot(as.data.frame(moranTest$res[c(1:999)]), aes(moranTest$res[c(1:999)])) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(aes(xintercept = moranTest$statistic), colour = "#FA7800",size=1) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title="Observed and permuted Moran's I",
       subtitle= "Observed Moran's I in orange",
       x="Moran's I",
       y="Count") +
  plotTheme()

```

## Etc.

Describe what we did here.

```{r Etc}
#Map of predicted value on both == 1 and == 0
test.Data <-
  test.Data %>%
  mutate(SalePrice.Predict = predict(reg1, test.Data))

test.Data.sf <-
  left_join(test.Data, studentData) %>%
  st_as_sf()

merge.predicted <-
  train.Data.2 %>%
  select(-SalePrice.Error,-SalePrice.AbsError, -SalePrice.APE)

merge.predicted <-
  rbind(merge.predicted,test.Data.sf)

ggplot()+
  geom_sf(data = BoulderCounty_Bundary, fill = "grey70") +
  geom_sf(data = BoulderMuni_Boundary, aes(fill = Municipality, alpha=0.5),colour = "white")+
  geom_sf(data = merge.predicted, aes(colour = q5(SalePrice.Predict)),size=.85)+
  scale_colour_manual(values = palette5,
                      labels=qbr(merge.predicted,"SalePrice.Predict"),
                      name = "Predicted Sale Price")+
  labs(title = "Predicted Sale Prices in Boulder County",
       subtitle = "Bouolder, Colorado ($)",
       caption = "Fig.2")+
  mapTheme()

#Adding Neighborhood Layer

boulder.nb <-
st_read("Boulder Neighborhoods.kml")%>%
  st_transform('ESRI:102254')%>%
  select(Name,geometry)

#Calculate AbsError and MAPE by neighborhoods
neighborhood_with_error <-
  st_join(boulder.nb,train.Data.2)

neighborhood_with_error <-
  neighborhood_with_error %>%
  group_by(Name)%>%
  summarize("Mean.absE" = mean(SalePrice.AbsError, na.rm = T),
            "MAPE" = mean(SalePrice.APE, na.rm = T))%>%
  mutate(MAPE.Pct = MAPE*100)

#Map of AbsError and Neighborhoods
ggplot()+
  geom_sf(data = BoulderCounty_Bundary, fill = "grey70") +
  geom_sf(data = neighborhood_with_error, aes(fill = MAPE.Pct),colour = "white")+
    scale_fill_gradient2(
      low = "red",
      mid = "white",
      high = "blue",
      midpoint = 0,
      space = "Lab",
      na.value = "grey50",
      guide = "colourbar",
      aesthetics = "fill",
    name = "Absolute Percentage Error" ) +
  mapTheme()

#Municipality with Error
muni_with_error <-
  st_join(BoulderMuni_Boundary,train.Data.2)%>%
  select(price,Municipality.x,SalePrice.APE,SalePrice.AbsError)

muni_with_error <-
  muni_with_error %>%
  group_by(Municipality.x)%>%
  summarize("Mean.absE" = mean(SalePrice.AbsError, na.rm = T),
            "MAPE" = mean(SalePrice.APE, na.rm = T),
            "Mean.price" = mean(price), na.rm = T)%>%
  mutate(MAPE.Pct = MAPE*100)

#Census Track with Error
tracts_with_error <-
  st_join(tracts19 ,train.Data.2)%>%
  select(price,GEOID.x,SalePrice.APE,SalePrice.AbsError)

tracts_with_error <-
  tracts_with_error %>%
  group_by(GEOID.x)%>%
  summarize("Mean.absE" = mean(SalePrice.AbsError, na.rm = T),
            "MAPE" = mean(SalePrice.APE, na.rm = T),
            "Mean.price" = mean(price), na.rm = T)%>%
  mutate(MAPE.Pct = MAPE*100)

ggplot()+
  geom_sf(data = BoulderCounty_Bundary, fill = "grey70") +
  geom_sf(data = tracts_with_error, aes(fill = MAPE.Pct),colour = "white")+
  scale_fill_gradient2(
    low = "red",
    mid = "white",
    high = "blue",
    midpoint = 0,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill",
    name = "Absolute Percentage Error" ) +
  mapTheme()

#scatterplot plot of MAPE by neighborhood as a function of mean price by neighborhood.

ggplot(data = tracts_with_error, aes(Mean.price, MAPE.Pct)) +
  geom_point(size = 1.5,colour = "orange") + 
  labs(x = "Mean Housing Prices",y = "Mean Absolute Percentage Errors", 
       title = "Absolute Percentage Error as sa function of Mean Housing Prices by Municipalities",
       subtitle = "Boulder County, CO")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.line = element_line(colour = "grey50", size = 1),
        panel.grid.major = element_line(linetype = "dotted",size = 1))

```

